---
title: 'ESM 244: Lab 5'
author: "Hanna Buechi"
date: "2/6/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages_data}

library(tidyverse)
library(tseries)
library(forecast)

energy <- read_csv("energy.csv")

```

c. Create 'ts' time series data using information in the res_total column from 'energy' data frame.

```{r}

res_ts <- ts(energy$res_total, frequency = 12, start = c(1973,1)) # 12 indicates monthly data, starts in January of 1973)
# View(res_ts) gives different output than res_ts

plot(res_ts)

# looks additive, intermediate peaks get larger through time
```

d. Decompose ts

```{r}

res_dc <- decompose(res_ts) # does Steps 1-4 of decomposition from class (ie. trend, seasonal, residuals)
plot(res_dc) # original, trend, seasonality, residuals that show trends that don't depend on the other decompositions

```

e. Other cool ways to present time-series data

```{r monthplot}

# a lot of information in a simple way

monthplot(res_ts) # over the years that have been studied, there has been a large spike in residential energy usage in summer months (that intermediate peak!)

```

```{r seasonplot}

ggseasonplot(res_ts) # similar pattern each year, more recent years are higher

ggseasonplot(res_ts, polar = TRUE) # what's the y-axis for August here?

```

f. Simple moving averages

forecast::ma()

```{r}

sma_res <- ma(res_ts, order = 5) # window size = 5, NAs on the ends means end value were cut because they don't have a window
plot(sma_res)

# play around with changing window size with plotting

plot(res_ts)
lines(sma_res, col = "red") # run simultaneously to get original data plot with moving average line

```

g. ACF (autocorrelation)

```{r}

res_acf <- acf(res_ts) # acf = autocorrelation function

# drastic peaks every 12 months with minor correlations on either side

ggtsdisplay(res_ts) # displays original data and ACF, which is nice

```


i. Holt-Winters exponential smoothing

```{r}

res_hw <- HoltWinters(res_ts)
res_hw

# additive (yep), trend (yep)
# alpha, beta, gamma for smoothing
# coefficients will be used in linear combinations for forecasting

plot(res_hw)

# how does model look with original data? (before predicting)

```

Now some future forecasting with Holt-Winters

```{r}


res_forecast <- forecast(res_hw, h = 60) # H-W model we already created 60 increments - 5 years - into the future
plot(res_forecast)

# recent slight downward or constant trend is dominating
# dark grey ribbons 80% CI, light grey 95% CI
# call res_forecast values into ggplot()

```

Check distribution of the residuals (good - I want to think about this more)

```{r}

par(mfrow = c(1,2))
hist(res_forecast$residuals)
qqnorm(res_forecast$residuals) # NICE


```


Autoregressive integrated moving average (ARIMA) using auto.arima() for p, d, q.

```{r}

res_pdq <- auto.arima(res_ts)
res_pdq

# (p,d,q) non-seasonal = (1,0,2), (P,D,Q) seasonal = (0,1,1)
```

(p,d,q) non-seasonal = (1,0,2), (P,D,Q) seasonal = (0,1,1)

Fit the ARIMA model using those pdq inputs:
```{r}

res_arima <- arima(res_ts, order = c(1,0,2), seasonal = list(order = c(0,1,1)))


# Check out the residuals

par(mfrow = c(1,2))
hist(res_arima$residuals)
qqnorm(res_arima$residuals)

# might be a little bit "leptokirtic" - sharper than normal

```

Now forecast with ARIMA outputs:

```{r}

forecast_res <- forecast(res_arima, h = 72)
plot(forecast_res)

# similar to H-W plot because there is good seasonality

```

Now let's try viewing this in ggplot:
```{r}

res_df <- data.frame(forecast_res) # coerce list to dataframe

# point forecast is forecasted estimate with 80% and 95% CI

month_seq <- seq(1,72)
res_df_2 <- data.frame(month_seq, res_df) # those row names....

ggplot(res_df_2, aes(x = month_seq, y = Point.Forecast)) +
  geom_line()

```















